---
title: "Create database of amphibian dermal exposure data"
---

```{r eval=FALSE, echo=FALSE}
#ctrl-shift k to knit r markdown
#https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf
```

Introduction
========================================================
Comparison of proposed amphibian dermal exposure model (Draft 8/19/15) with collected dermal exposure data. Data set is from Van Meter et al. 2014 and Van Meter et al. 2015.

Computational environment
========================================================
Location of this repository: 
https://github.com/puruckertom/amphib_data
```{r eval=FALSE, echo=FALSE}
#library(installr) #windows only currently
#updateR()
```
Version and installed libraries.
```{r eval=TRUE, echo=TRUE}
R.Version()$version.string
Sys.info()[4]
```

```{r eval=TRUE, echo=FALSE}
#tom computer
if(Sys.info()[4]=="DC2626UTPURUCKE"){
  amphibdir <- "C:\\git\\VanMeteretal2015_AECT\\data\\"
}

# marcia mac
if(Sys.info()[4] == "Marcias-MacBook-Pro.local"){
  amphibdir <- path.expand("~/git/VanMeteretal2015_AECT/data/")
}

# marcia mac
if(Sys.info()[4] == "Marcias-MacBook-Pro.local"){
  frogsoildir <- path.expand("~/git/VanMeteretal2016_ple_v_ols/")
}

#tom computer
if(Sys.info()[4]=="DZ2626UTPURUCKE"){
  frogsoildir <- "C:\\git\\VanMeteretal2016_ple_v_ols\\"
}
  
# marcia mac
if(Sys.info()[4] == "Marcias-MacBook-Pro.local"){
  amphibdatadir <- path.expand("~/git/amphib_data/")
}
  
#tom computer
if(Sys.info()[4]=="DZ2626UTPURUCKE"){
  amphibdatadir <- "C:\\git\\amphib_data\\"
}
```

Import data from consolidated database
```{r eval=TRUE, echo=FALSE}
##################
#the data set
##################
# import data 
combined_data <- read.table(paste(amphibdatadir,"combined_data.csv",sep=""), header = TRUE, sep = ",") 
```

Models from dust_check.R
``` {r eval=TRUE, echo=FALSE}
#### X1 foliar spray exposure functions for bird and mammal ####
X1_new_bird<-function(a,bw2){
  x1_bird<- ((a*(10*bw2^0.667) * 6.01*6*0.079*0.2) + (a*(10*bw2^0.667) * 0.1)) / (bw2 /1000)                 
  return(x1_bird)
}

X1_new_mamm<-function(a,bw2){
  x1_mamm<- ((a*(12.3*bw2^0.65) * 6.01*6*0.079*0.2) + (a*(12.3*bw2^0.65) * 0.1)) / (bw2 /1000)                 
  return(x1_mamm)
}

#### X3 bare ground exposure functions for bird and mammal ####
# constants
# assumed depth of soil = 1 (cm)
# assumed soil density = 1300 (mg soil/cm3)
# soil to skin adherence factor (upper bound human estimate of gardner's hands) = 0.958 (mg/cm2)

x3_bird<-function(a,bw2){
  x3<-(((a*1/(1*1300))*0.958*(10*bw2^0.667))+(a*(10*bw2^0.667)*0.1)) / (bw2 /1000)
  return (x3)
}
x3_mamm<-function(a,bw2){
  x3<-(((a*1/(1*1300))*0.958*(12.3*bw2^0.65))+(a*(12.3*bw2^0.65)*0.1)) / (bw2 /1000)
  return (x3)
}
```

Proposed dermal exposure model parameters for amphibians (Document 8/19/15):

Dermal dose through direct spray-
**** Equation 12. ****
Dermal spray dose = [(Arate * 11.2) (SAtotal * 0.5) * DAF / BW ]* Fred

Arate = application rate (lb a.i./A)
11.2 = conversion from lb a.i./A to ug a.i./g-bw
SAtotal = Total surface area of organism
0.5 = Assumed that either the top half or the bottom half of the animal will be in contact with the ground or direct spray
DAF = dermal adsorption fraction used to account for pesticide specific data that define a fraction of the pesticide mass present on the animal that is actually absorbed
BW = body weight
Fred = Dermal route equivalency factor applied to dermal exposure to derive an estimate of oral dose. Assumed to be 1 for amphibians.
SAtotal = Surface area total = a3 * BW^b3
a3 = 1.131
b3= 0.579

Same surface area formula used by Van Meter et al. 

Dermal exposure function for dermal spray dose
``` {r eval=TRUE, echo=FALSE}
dermal_spray_dose<-function(Arate, SAtotal, DAF, BW, Fred){
  dsd1<-(((Arate * 11.2) * (SAtotal * 0.5) * DAF / BW )* Fred)
  return (dsd1)
}
#test function returns a value
dermal_spray_dose(7,8,9,10,11)
```

Dermal dose through foliage contact-
**** Equation 15. **** 
Dermal contact dose = ( Cplant(t) * Fdfr * Rfoliar contact ^ (8*SAtotal*0.079*0.1)     / BW ) * Fred

Cplant = Concentration of the pesticide in crop foliage at time t. Residue value used for broadleaf foliage concentration in the assessment of dietary exposure
Fdfr = Dislodgeable foliar residue adjustment factor
Rfoliar contact = Rate of foliar contact
8 = number of hours per day the animal is active
SAtotal = surface area total of the organism
0.079 = represents the fraction of the animal in contact with the foliage
0.1 = unit conversion to ug a.i./g-bw
BW = body weight of organism
Fred = Dermal route equivalency factor (1 for amphibians)

Dermal exposure function for dermal dose through foliage contact
``` {r eval=TRUE, echo=FALSE}
dermal_contact_dose<-function(Cplant,Fdfr,Rfoliarcontact,SAtotal,BW,Fred){
  dcd1<-( Cplant * Fdfr * Rfoliarcontact ^ (8*SAtotal*0.079*0.1)     / BW ) * Fred
  return (dcd1)
}
#test function returns a value
dermal_contact_dose (1,2,3,4,5,6)
```

Use input values from Van Meter et al. database to get estimated dermal spray dose and compare with measured body burdes
``` {r eval=TRUE, echo=FALSE}
combined_data$sa <- 1.131 * (combined_data$bodyweight ^ 0.579)
DAF = 1
Fred = 1
dermal_spray_dose_out<-dermal_spray_dose(combined_data$app_rate_g_cm2, combined_data$sa, DAF, combined_data$bodyweight, Fred)

```

Use input values from Van Meter et al. database to get estimated dermal dose through foliar contact and compare with measured body burdes
``` {r eval=TRUE, echo=FALSE}
dermal_spray_dose_out<-dermal_spray_dose()
```


