---
title: "Create database of amphibian dermal exposure data"
---

```{r eval=FALSE, echo=FALSE}
#ctrl-shift k to knit r markdown
#https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf
```

Introduction
========================================================
Comparison of proposed amphibian dermal exposure model (Draft 8/19/15) with collected dermal exposure data. Data set is from Van Meter et al. 2014 and Van Meter et al. 2015.

Computational environment
========================================================
Location of this repository: 
https://github.com/puruckertom/amphib_data
```{r eval=FALSE, echo=FALSE}
#library(installr) #windows only currently
#updateR()
```
Version and installed libraries.
```{r eval=TRUE, echo=TRUE}
R.Version()$version.string
Sys.info()[4]
```

```{r eval=TRUE, echo=FALSE}
#tom computer
if(Sys.info()[4]=="DC2626UTPURUCKE"){
  amphibdir <- "C:\\git\\VanMeteretal2015_AECT\\data\\"
}

# marcia mac
if(Sys.info()[4] == "Marcias-MacBook-Pro.local"){
  amphibdir <- path.expand("~/git/VanMeteretal2015_AECT/data/")
}

# marcia mac
if(Sys.info()[4] == "Marcias-MacBook-Pro.local"){
  frogsoildir <- path.expand("~/git/VanMeteretal2016_ple_v_ols/")
}

#tom computer
if(Sys.info()[4]=="DZ2626UTPURUCKE"){
  frogsoildir <- "C:\\git\\VanMeteretal2016_ple_v_ols\\"
}
  
# marcia mac
if(Sys.info()[4] == "Marcias-MacBook-Pro.local"){
  amphibdatadir <- path.expand("~/git/amphib_data/")
}
  
#tom computer
if(Sys.info()[4]=="DZ2626UTPURUCKE"){
  amphibdatadir <- "C:\\git\\amphib_data\\"
}
```

Import data from consolidated database
```{r eval=TRUE, echo=FALSE}
##################
#the data set
##################
# import data 
combined_data <- read.table(paste(amphibdatadir,"combined_data.csv",sep=""), header = TRUE, sep = ",") 
```

Proposed dermal exposure model parameters for amphibians (Document 8/19/15):

Dermal dose through direct spray-
**** Equation 12. ****
Dermal spray dose = [(Arate * 11.2) (SAtotal * 0.5) * DAF / BW ]* Fred

Arate = application rate (lb a.i./A)
11.2 = conversion from lb a.i./A to ug a.i./g-bw
SAtotal = Total surface area of organism
0.5 = Assumed that either the top half or the bottom half of the animal will be in contact with the ground or direct spray
DAF = dermal adsorption fraction used to account for pesticide specific data that define a fraction of the pesticide mass present on the animal that is actually absorbed
BW = body weight
Fred = Dermal route equivalency factor applied to dermal exposure to derive an estimate of oral dose. Assumed to be 1 for amphibians.
SAtotal = Surface area total = a3 * BW^b3
a3 = 1.131
b3= 0.579

Same surface area formula used by Van Meter et al. 

Dermal exposure function for dermal spray dose
``` {r eval=TRUE, echo=FALSE}
#convert from g/cm2 to lbs/Acre for input into dermal spray dose function
convert_app_rate<-function(app_rate){
  new_app_rate<-(app_rate * 89217.9122)
  return(new_app_rate)
}
#function to calculate dermal direct spray dose
dermal_spray_dose<-function(Arate, SAtotal, DAF, BW, Fred){
  dsd1<-(((Arate * 11.2) * (SAtotal * 0.5) * DAF / BW )* Fred)
  return (dsd1)
}
#test function returns a value
convert_app_rate(8)
dermal_spray_dose(7,8,9,10,11)
```

Dermal dose through foliage contact-
**** Equation 15. **** 
Dermal contact foliar dose = ( Cplant(t) * Fdfr * Rfoliar contact ^ (8*SAtotal*0.079*0.1)     / BW ) * Fred

Cplant = Concentration of the pesticide in crop foliage at time t. Residue value used for broadleaf foliage concentration in the assessment of dietary exposure
For broadleaf plants Cplant = 135 * application rate (lbs a.i./hectare)
Fdfr = Dislodgeable foliar residue adjustment factor
Rfoliar contact = Rate of foliar contact
8 = number of hours per day the animal is active
SAtotal = surface area total of the organism
0.079 = represents the fraction of the animal in contact with the foliage
0.1 = unit conversion to ug a.i./g-bw
BW = body weight of organism
Fred = Dermal route equivalency factor (1 for amphibians)

Dermal exposure function for indirect dermal dose through foliage contact
``` {r eval=TRUE, echo=FALSE}
# 220462 = unit conversion from g/cm2 to lbs/hectare
# 135 = used in T-Rex to calculate residue on broadleaf plant
# function to calculate Cplant (broadleaf foliar residue from T-Rex)
Cplant_calc<-function(app_rate_g_cm2){
  Cplant_out<-(app_rate_g_cm2 * 220462 * 135)
  return(Cplant_out)
  }
# function to calculate dermal exposure through indirect foliar contact
dermal_contact_dose<-function(Cplant,Fdfr,Rfoliarcontact,SAtotal,BW,Fred){
  dcd1<-( Cplant * Fdfr * Rfoliarcontact ^ (8*SAtotal*0.079*0.1)     / BW ) * Fred
  return (dcd1)
}
#test function returns a value
out1<-Cplant_calc(4)
dermal_contact_dose (out1,2,3,4,5,6)
```

Use input values from Van Meter et al. database to get estimated dermal spray dose through direct spray and compare with measured body burdens
``` {r eval=TRUE, echo=FALSE}
# surface area calculation
combined_data$sa <- 1.131 * (combined_data$bodyweight ^ 0.579)
# direct dermal spray dose estimated
DAF = 1 #constant
Fred = 1 #constant
combined_data$app_rate_lbs_A<-convert_app_rate(combined_data$app_rate_g_cm)
combined_data$dermal_spray_dose_out<-dermal_spray_dose(combined_data$app_rate_lbs_A, combined_data$sa, DAF, combined_data$bodyweight, Fred)
combined_data$ratio_out<-combined_data$dermal_spray_dose_out / combined_data$TissueConc
```
Use input values from Van Meter et al. database to get estimated dermal spray dose through indirect foliage and compare with measured body burdens
``` {r eval=TRUE, echo=FALSE}
# surface area calculation
combined_data$sa <- 1.131 * (combined_data$bodyweight ^ 0.579)
# DAF = 1 #constant
# Fred = 1 #constant
# Fdfr = 0.62
# Rfoliarcontact = 6.01

combined_data$Cplant<-Cplant_calc(combined_data$app_rate_g_cm2)
combined_data$dermal_foliage_dose_out<-dermal_contact_dose(combined_data$Cplant, 0.62, 6.01, combined_data$sa, combined_data$bodyweight, 1)
combined_data$ratio_out_foliage<-combined_data$dermal_foliage_dose_out / combined_data$TissueConc
```

Plots comparing modeled with measured endpoints for direct vs. direct and indirect vs. indirect
```{r fig.width=7, fig.height=6}
#plots for direct spray dose equation
max1<-max(combined_data$TissueConc, na.rm=TRUE)
min<-0
plot(combined_data$dermal_spray_dose_out~combined_data$TissueConc, xlim=c(min,max1), ylim=c(min,max1), xlab="Measured body burden", ylab="DUST direct spray")
abline(0, 1, col="red")
title(main=c("Direct spray with all body burden data"))
#histogram of ratios
hist(combined_data$ratio_out)
#plots for indirect foliage dose 
max1<-max(combined_data$dermal_foliage_dose_out, na.rm=TRUE)
min<-0
plot(combined_data$dermal_foliage_dose_out~combined_data$TissueConc, xlim=c(min,max1), ylim=c(min,max1), xlab="Measured body burden", ylab="DUST indirect foliage")
abline(0, 1, col="red")
title(main=c("Indirect foliage spray with all body burden data"))

hist(combined_data$ratio_out_foliage)
```

Plot by chemical modeled with measured endpoints for all comparisons
```{r fig.width=7, fig.height=6}
unique.chemicals<-unique(combined_data$Chemical)
# create consistent chemical names
combined_data$Chemical[combined_data$Chemical=="ATZ"]<-"Atrazine"
combined_data$Chemical[combined_data$Chemical=="Imid"]<-"Imidacloprid"
combined_data$Chemical[combined_data$Chemical=="Fip"]<-"Fipronil"
combined_data$Chemical[combined_data$Chemical=="Pendi"]<-"Pendimethalin"
combined_data$Chemical[combined_data$Chemical=="TDN"]<-"Triadimefon"
chemicals2<-c("Atrazine","Imidacloprid","Fipronil","Pendimethalin","Triadimefon")


par(mfrow=c(1,1))
# direct dermal DUST vs. measured direct
max1<-max(combined_data$TissueConc)
min<-0 
plot(combined_data$dermal_spray_dose_out[which(combined_data$Application=="Direct")]~combined_data$TissueConc[which(combined_data$Application=="Direct")], xlab=c('Measured Tissue Concentration Direct'), ylab=c("Modeled Tissue Concentration Direct"),asp=1, pty='s', xlim=c(min,max1), ylim=c(min,max1),pch=16, col=combined_data$Species[which(combined_data$Application=="Direct")])
abline(0, 1, col="red")
title(main=c("Direct dermal DUST vs. measured direct"))

# direct dermal DUST vs. measured indirect
max1<-max(combined_data$TissueConc)
min<-0 
plot(combined_data$dermal_spray_dose_out[which(combined_data$Application=="Indirect")]~combined_data$TissueConc[which(combined_data$Application=="Indirect")], xlab=c('Measured Tissue Concentration Indirect'), ylab=c("Modeled Tissue Concentration Direct"),asp=1, pty='s', xlim=c(min,max1), ylim=c(min,max1), pch=16,col=combined_data$Species[which(combined_data$Application=="Indirect")])
abline(0, 1, col="red")
title(main=c("Direct dermal DUST vs. measured indirect"))

# indirect dermal DUST vs. measured indirect
max1<-max(combined_data$TissueConc)
min<-0 
plot(combined_data$dermal_foliage_dose_out[which(combined_data$Application=="Indirect")]~combined_data$TissueConc[which(combined_data$Application=="Indirect")], xlab=c('Measured Tissue Concentration Indirect'), ylab=c("Modeled Tissue Concentration Indirect"),asp=1, pty='s', xlim=c(min,max1), ylim=c(min,max1), pch=16,col=combined_data$Species[which(combined_data$Application=="Indirect")])
abline(0, 1, col="red")
title(main=c("Indirect dermal DUST vs. measured indirect"))

# indirect dermal DUST vs. measured direct
max1<-max(combined_data$TissueConc[which(combined_data$Application=="Direct")])
min<-0 
plot(combined_data$dermal_foliage_dose_out[which(combined_data$Application=="Direct")]~combined_data$TissueConc[which(combined_data$Application=="Direct")], xlab=c('Measured Tissue Concentration Direct'), ylab=c("Modeled Tissue Concentration Indirect"),asp=1, pty='s', xlim=c(min,max1), ylim=c(min,max1), pch=16,col=combined_data$Species[which(combined_data$Application=="Direct")])
abline(0, 1, col="red")
title(main=c("Indirect dermal DUST vs. measured direct"))

```


