---
title: "Compare Amphibian Dermal Exposure Data with Proposed Models"
author: "Purucker ST, Snyder MN, Glinski DA, Van Meter RJ, Garber K, Cyterski MJ, Sinnathamby S, Henderson WM"
---

```{r eval=FALSE, echo=FALSE}
#ctrl-shift k to knit r markdown
#https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf
```

<br><br>

***

**Overview**
======================================================================
A comparison between the Terrestrial Investigation Model (TIM) and the collected amphibian dermal exposure data (introduced in 00_amphib_data_combine.Rmd) will be conducted.

Collected amphibian dermal exposure data is from 9 studies. Van Meter et al. 2014/2015 data includes exposure with 5 active ingredients (imidacloprid, pendimethalin, atrazine, fipronil, and triadimefon) and 9 amphibian species in the terrestrial metamorph stage (American toad, Barking treefrog, Cricket frog, Fowler's toad, Gray treefrog, Green treefrog, Leopard frog, Mole salamander, Narrowmouth toad). Van Meter et al. 2016 data includes exposure to the same 5 pesticides on American toads. Van Meter et al. 2018 data consisted of exposure to a single pesticide or pesticide mixtures (atrazine, metolachlor, 2,4-D, malathion, propiconazole) on juvenile green frogs. Glinski et al. 2018a (Dehydration) data includes exposure to 5 pesticides (atrazine, triadimefon, metolachlor, chlorothalonil, imidacloprid) on 2 amphibian species (Southern leopard frogs, Fowler's toads). Glinski et al. 2018b (Metabolites) data consists of exposure to 3 pesticides (atrazine, triadimefon, fipronil) on Fowler's toads. Glinski et al. 2018c (Biomarkers) data includes exposure to single, double, or triple pesticide mixtures (bifenthrin, metolachlor, triadimefon) on Southern leopard frogs.  Glinski et al. 2020 (Dermal Routes) data includes exposure to bifenthrin, chlorpyrifos, and trifloxystrobin on Leopard frogs. Henson-Ramsey 2008 data consists of exposure to malathion for Tiger salamanders.


```{r eval = TRUE, echo = FALSE, include = FALSE}
# Install required libraries
library(ggplot2)
require(gridExtra)
library(reshape2)


# tom epa linux
if(Sys.info()[4] == "d2626ut7920d.rtpnc.epa.gov"){
  amphibdir <- path.expand("~/git/amphib_dermal_collation/")
}
# Emma's EPA
if(Sys.info()[4]=="LZ2626UECHELSVI"){
  amphibdir <- path.expand("C:/Users/echelsvi/git/amphib_dermal_collation/")
}

# Variable names for data
amphibdir_data_in <- paste(amphibdir,'data_in/',sep='')
amphibdir_data_out <- paste(amphibdir,'data_out/',sep='')
amphibdir_graphics <- paste(amphibdir,'graphics/',sep='')
amphibdir_src <- paste(amphibdir,'src/',sep='')
```
<br><br>

***

**Collated Data Set**
============================================================================
**Data Set Dimensions, Column Names, and Summary:**
```{r eval = TRUE, echo = FALSE}
##################
#the data set
##################
# import data from consolidated database
combined_data_filename <- paste(amphibdir_data_out,"amphib_dermal_collated.csv", sep="")
combined_data <- read.csv(combined_data_filename)
dim(combined_data)
colnames(combined_data)
summary(combined_data)
```


~~ DELETE THIS ONCE GET APP RATES FOR DERMAL ROUTES DATA!!
```{r eval = TRUE, echo = TRUE}
combined_data <- combined_data[!(combined_data$source == "dag_dermal_routes"),]
```
<br><br>

***

**TIM (Terrestrial Investigation Model)**
=====================================================================
By default, TIM estimates exposures for birds. The dermal exposure dose (body burden) is computed by combining the pesticide application rate (app_rate), surface area (sa_default_tim), fraction of surface area exposed (sa_tim_default_frac), body weight of the exposed organism (body_weight) and the dermal absorption factor (dermal_af).  


$$TIM = \frac{AR * SA * SAF * DAF}{BW}$$

```{r eval = TRUE, echo = FALSE, include = FALSE}
colnames(combined_data)
n <- length(combined_data$tissue_conc_ugg)

# bird model with perez body weight
#   we dont care about perez body weights because they are based on large pigeons
# sa_bird_perez <- 65.15 * (body_weight ^ 0.559)
```

***

### **TIM Default Calculation**
```{r eval = TRUE, echo = TRUE}
# Equation 1: TIM Default Values 
app_rate <- combined_data$app_rate_g_cm2
body_weight <- combined_data$body_weight_g
conv_rate <- 1000000.0
dermal_af <- 1.0
hours_one <- 1.0
sa_default_tim <- 10 * (body_weight ^ 0.667) 
sa_tim_default_frac <- 0.5
tissue_conc <- combined_data$tissue_conc_ugg
```

```{r eval = TRUE, echo = TRUE, include = TRUE}
# Calculate Equation 1a: TIM Default
combined_data$tim_direct_default <- (app_rate * conv_rate * sa_default_tim * sa_tim_default_frac * dermal_af)/body_weight
tim_direct_default <- combined_data$tim_direct_default
```

```{r eval = TRUE, echo = FALSE, include = TRUE}
hist(tim_direct_default, main = "TIM Default", xlab = "Dermal Exposure Dose", col = "#636363")
```


~~~~ MAYBE DELETE THIS OR PUT IN APPENDIX AT BOTTOM UNLESS USEFUL HERE/ELSEWHERE
```{r eval = TRUE, echo = FALSE, include = TRUE}
layout(matrix(c(1,2,3,4), nrow = 2, ncol = 2, byrow= TRUE)) #set the plotting area into a 2row*2col array
hist(combined_data$tim_direct_default[combined_data$source == "rvm2015"], xlab = "Dermal Exposure Dose", main = "Van Meter 2014/15", col = "#636363")
hist(combined_data$tim_direct_default[combined_data$source == "rvm2016"], xlab = "Dermal Exposure Dose", main = "Van Meter 2016", col = "#636363")
hist(combined_data$tim_direct_default[combined_data$source == "rvm2017"], xlab = "Dermal Exposure Dose", main = "Van Meter 2017", col = "#636363")
hist(combined_data$tim_direct_default[combined_data$source == "hr2008"], xlab = "Dermal Exposure Dose", main = "Henson-Ramsey 2008", col = "#636363")


layout(matrix(c(1,2,3,3), nrow = 2, ncol = 2, byrow= TRUE)) #set the plotting area into a 2row*2col array
hist(combined_data$tim_direct_default[combined_data$source == "dag_dehydration"], xlab = "Dermal Exposure Dose", main = "Glinski (Dehydration)", col = "#636363")
hist(combined_data$tim_direct_default[combined_data$source == "dag_metabolites"], xlab = "Dermal Exposure Dose", main = "Glinski (Metabolites)", col = "#636363")
hist(combined_data$tim_direct_default[combined_data$source == "dag_biomarker"], xlab = "Dermal Exposure Dose", main = "Glinski (Biomarkers)", col = "#636363")
#hist(combined_data$tim_direct_default[combined_data$source == "dag_dermal_routes"], xlab = "Dermal Exposure Dose", main = "Glinski (Dermal Routes)", col = "#636363")
```


### **TIM Default Ratios**
~~ blurb about how we can assess the ratios and talk about below scatterplots...1:1 would be perfect agreement
```{r eval = TRUE, echo = TRUE}
# Calculate Equation 1b: TIM Default Ratios
combined_data$tim_direct_default_ratios <- combined_data$tim_direct_default/tissue_conc
```

~~ talk about this range in values (or possibly exclude if it isn't helpful?)
```{r eval = TRUE, echo = TRUE, include = TRUE}
# Range
print(paste(min(combined_data$tim_direct_default_ratios), "-",   max(combined_data$tim_direct_default_ratios)))
```


```{r eval = TRUE, echo = FALSE, include = TRUE}
ggplot(combined_data, aes(x = tissue_conc, y = tim_direct_default, color = combined_data$source)) + 
  geom_point() + 
  labs(title = "Modeled vs. Measured, by Study", x = "Measured Tissue Concentrations", y = "TIM Default") +
  geom_abline(intercept = 0, slope = 1) + 
  theme(legend.position = "bottom", legend.title = element_blank())
```

```{r eval = TRUE, echo = FALSE, include = TRUE}
ggplot(combined_data, aes(x = tissue_conc, y = tim_direct_default, color = combined_data$chemical)) + 
  geom_point() + 
  labs(title = "Modeled vs. Measured, by Chemical", x = "Measured Tissue Concentrations", y = "TIM Default") +
  geom_abline(intercept = 0, slope = 1) +
  theme(legend.position = "bottom", legend.title = element_blank())
```

```{r eval = TRUE, echo = FALSE, include = TRUE}
ggplot(combined_data, aes(x = tissue_conc, y = tim_direct_default, color = combined_data$application)) + 
  geom_point() + 
  labs(title = "Modeled vs. Measured, by Application Type", x = "Measured Tissue Concentrations", y = "TIM Default") +
  geom_abline(intercept = 0, slope = 1) +
  theme(legend.position = "bottom", legend.title = element_blank())
```


***

### **TIM Amphibian Calculation**
To gain a better understanding of pesticide risks on amphibians, the dermal exposure dose was also computed with a modified allometric relationship between body weight and surface area that is better suited for amphibians (Hutchison et al. 1968, EPA 2009).
```{r eval = TRUE, echo = TRUE, include = TRUE}
# Equation 2: TIM Amphibians value (Hutchinson body weight)
sa_amphib_hutchinson <- 1.131 * (body_weight ^ 0.579)
```

```{r eval = TRUE, echo = FALSE}
# sa_amphib_frac <- 2.0
```

```{r eval = TRUE, echo = TRUE, include = TRUE}
# Calculate Equation 2a: TIM Amphibian
combined_data$tim_direct_amphib <- (app_rate * conv_rate * sa_amphib_hutchinson * sa_tim_default_frac * dermal_af)/body_weight

tim_direct_amphib <- combined_data$tim_direct_amphib
```


```{r eval = TRUE, echo = FALSE, include = TRUE}
hist(tim_direct_amphib, main = "TIM Amphibian", xlab = "Dermal Exposure Dose", col = "#636363")
```


~~~~ MAYBE DELETE THIS OR PUT IN APPENDIX AT BOTTOM UNLESS USEFUL HERE/ELSEWHERE
```{r eval = TRUE, echo = FALSE, include = TRUE}
layout(matrix(c(1,2,3,4), nrow = 2, ncol = 2, byrow= TRUE)) #set the plotting area into a 2row*2col array
hist(combined_data$tim_direct_amphib[combined_data$source == "rvm2015"], xlab = "Dermal Exposure Dose", main = "Van Meter 2014/15", col = "#636363")
hist(combined_data$tim_direct_amphib[combined_data$source == "rvm2016"], xlab = "Dermal Exposure Dose", main = "Van Meter 2016", col = "#636363")
hist(combined_data$tim_direct_amphib[combined_data$source == "rvm2017"], xlab = "Dermal Exposure Dose", main = "Van Meter 2017", col = "#636363")
hist(combined_data$tim_direct_amphib[combined_data$source == "hr2008"], xlab = "Dermal Exposure Dose", main = "Henson-Ramsey 2008", col = "#636363")


layout(matrix(c(1,2,3,3), nrow = 2, ncol = 2, byrow= TRUE)) #set the plotting area into a 2row*2col array
hist(combined_data$tim_direct_amphib[combined_data$source == "dag_dehydration"], xlab = "Dermal Exposure Dose", main = "Glinski (Dehydration)", col = "#636363")
hist(combined_data$tim_direct_amphib[combined_data$source == "dag_metabolites"], xlab = "Dermal Exposure Dose", main = "Glinski (Metabolites)", col = "#636363")
hist(combined_data$tim_direct_amphib[combined_data$source == "dag_biomarker"], xlab = "Dermal Exposure Dose", main = "Glinski (Biomarkers)", col = "#636363")
#hist(combined_data$tim_direct_amphib[combined_data$source == "dag_dermal_routes"], xlab = "Dermal Exposure Dose", main = "Glinski (Dermal Routes)", col = "#636363")
```


### **TIM Amphibian Ratios**
~~~ talk about briefly ... A model along the 1:1 line would display perfect agreement with the measured values.  
```{r eval = TRUE, echo = TRUE, include = TRUE}
# Calculate Equation 2b: TIM Amphibian Ratios
combined_data$tim_direct_amphib_ratios <- combined_data$tim_direct_amphib/tissue_conc
```

~~ need to include this range?
```{r eval = TRUE, echo = TRUE, include = TRUE}
# Range
print(paste(min(combined_data$tim_direct_amphib_ratios), "-",   max(combined_data$tim_direct_amphib_ratios)))
```


```{r eval = TRUE, echo = FALSE, include = TRUE}
ggplot(combined_data, aes(x = tissue_conc, y = tim_direct_amphib, color = combined_data$source)) + 
  geom_point() + 
  labs(title = "Modeled vs. Measured, by Study", x = "Measured Tissue Concentrations", y = "TIM Amphibian") +
  geom_abline(intercept = 0, slope = 1) +
  theme(legend.position = "bottom", legend.title = element_blank())
```

```{r eval = TRUE, echo = FALSE, include = TRUE}
ggplot(combined_data, aes(x = tissue_conc, y = tim_direct_amphib, color = combined_data$chemical)) + 
  geom_point() + 
  labs(title = "Modeled vs. Measured, by Chemical", x = "Measured Tissue Concentrations", y = "TIM Amphibian") +
  geom_abline(intercept = 0, slope = 1) +
  theme(legend.position = "bottom", legend.title = element_blank())
```

```{r eval = TRUE, echo = FALSE, include = TRUE}
ggplot(combined_data, aes(x = tissue_conc, y = tim_direct_amphib, color = combined_data$application)) + 
  geom_point() + 
  labs(title = "Modeled vs. Measured, by Application Type", x = "Measured Tissue Concentrations", y = "TIM Amphibian") +
  geom_abline(intercept = 0, slope = 1) +
  theme(legend.position = "bottom", legend.title = element_blank())
```
<br><br>

***

**Assessing Model Performance**
=======================================================
The TIM Default and TIM Amphibian models, described and shown above, were compared to one another and to the post-exposure amphibian body burdens found in the collated data set in order to assess model performance.


```{r eval = TRUE, echo = TRUE, include = TRUE}
# Create df of ratios
model_log_ratios <- as.vector(c(log10(combined_data$tim_direct_default_ratios), 
                        y=log10(combined_data$tim_direct_amphib_ratios)))
model_factor <- as.vector(c(rep('tim_default_direct',n),rep('tim_amphib_direct',n)))
df.ratios <- data.frame(x=model_factor, y=model_log_ratios)

# Create df of tissue_conc and TIM Amphibian
df.tim_amphib <- data.frame(x=tissue_conc, y=tim_direct_amphib)

# Create df of tissue_conc and TIM Default
df.tim_default <- data.frame(x=tissue_conc, y=tim_direct_default)
```

***

### **Comparing Models**
```{r eval = TRUE, echo = FALSE, include = TRUE}
ggplot(df.ratios, aes(x=x, y=y)) + 
  geom_point() + 
  labs(title="TIM Amphibian vs. TIM Default",  x="Model Factor", y = "TIM Value")
```

***

### **Plotting Modeled vs. Measured**

Scatter plots were utilized to assess the models. A model along the 1:1 line would display perfect agreement with the measured values.  


```{r eval = TRUE, echo = FALSE, include = TRUE}
plot_default <- ggplot(df.tim_default, aes(x=x, y=y)) + 
  geom_point() + 
  labs(title="Modeled vs. Measured",  x="Measured Tissue Concentrations", y = "TIM Default") + 
  geom_abline(intercept = 0, slope = 1)

plot_amphib <- ggplot(df.tim_amphib, aes(x=x, y=y)) +
  geom_point() + 
  labs(title="Modeled vs. Measured",  x="Measured Tissue Concentrations", y = "TIM Amphibian ") +
  geom_abline(intercept = 0, slope = 1)

grid.arrange(plot_default, plot_amphib, ncol = 2)
```

***

### **Plotting log10 Ratio of Modeled Burdens to Observed Burdens**

Box plots were utilized to further assess model performance. Well-performing models would minimize error, and therefore would have a higher percentage of ratios near zero (log10 of unity).  


```{r eval = TRUE, echo = FALSE, include = TRUE}
fill <- "#969696"
line <- "#252525"
ggplot(df.ratios, aes(x=x,y=y)) +
    geom_boxplot(fill = fill, colour = line) + 
    scale_x_discrete(name = "Model") +
    scale_y_continuous(name = "log10 of Model-Observation Ratio") +
    theme_bw()
model_boxplot_filename <- paste(amphibdir_graphics,"model_boxplot.png",sep='')
ggsave(model_boxplot_filename, device="png", width=4, height=4)
```

***

### **False Negative Rates**

Summary statistics for false negative rates were also computed for both models. False negative rates are defined as an outcome where the modeled value is less than the measured value. This rate is based on a Type II Error in statistical hypothesis testing. False negative rates assess how often a model under-estimates the measurement of interest.

#### **TIM Default**

~~ update summary (below) after all data updated/included

As displayed below, there are zero false negative rates for TIM Default. But,  115/798 (14.4%) of the measurements are within an order of magnitude of the modeled value. 
```{r eval = TRUE, echo = TRUE, include = TRUE}
# False Negative Rate: TIM Default
sum(tissue_conc > tim_direct_default)

# False Negative Rate (x10): TIM Default
sum(tissue_conc*10 > tim_direct_default)
```

#### **TIM Amphibian**
~~ update summary (below) after all data updated/included  

There are 107/798 (13.4%) false negative screen results for TIM Amphibian, and 576/798 (72.2%) measurements within an order of magnitude of the modeled value. TIM Amphibian uses a less conservative but more representative surface area calculation for amphibians. 
```{r eval = TRUE, echo = TRUE, include = TRUE}
# False Negative Rate: TIM Amphibian
sum(tissue_conc > combined_data$tim_direct_amphib)

# False Negative Rate (x10): TIM Amphibian
sum(tissue_conc*10 > combined_data$tim_direct_amphib)
```

***

### **Comparing Surface Area Calculations** 

While TIM Amphibian had a higher percentage of false negative screening values than TIM Default, the difference in screening values can be attributed to the surface area calculation. Amphibian surface area, as compared to the TIM Default surface area based on birds, is extrapolated from measured body weights. The comparison between surface area calculations is depicted below.    

```{r eval = TRUE, echo = FALSE}
min_bw <- min(combined_data$body_weight_g, na.rm=T)
max_bw <- max(combined_data$body_weight_g, na.rm=T)
range_bw <- seq(min_bw, max_bw, by = 0.01)
default_sa <- 10*range_bw^0.667
hutchison_sa <- 1.131 * range_bw^0.579
#max(default_sa/hutchison_sa)
sa_data <- data.frame(range_bw, default_sa, hutchison_sa)
ggplot(sa_data, aes(range_bw)) + 
  labs(x = "body weight (g)", y = "Surface Area (cm^2)",color = "Method") +
  geom_line(aes(y = default_sa, colour = "Default")) + 
  geom_line(aes(y = hutchison_sa, colour = "Hutchison")) +
  theme_bw() +
  theme(legend.position =c(0.2,0.8))
  
sa_comparison_filename <- paste(amphibdir_graphics,"sa_comparison.png",sep='')
ggsave(sa_comparison_filename, device="png", width=4, height=4)

```
<br><br>

***

Session Information
```{r eval = TRUE, echo = FALSE}
sessionInfo()
```