---
title: "Compare Amphibian Dermal Exposure Data with Proposed Models"
author: "Purucker ST, Snyder MN, Glinski DA, Van Meter RJ, Garber K, Cyterski MJ, Sinnathamby S, Henderson WM"
---

```{r eval=FALSE, echo=FALSE}
#ctrl-shift k to knit r markdown
#https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf
```
<br><br>

***

**Overview**
======================================================================
A comparison between the proposed amphibian dermal exposure model (DUST) from Draft 8/19/15 and the collected amphibian dermal exposure data (introduced in 00_amphib_data_combine.Rmd) will be conducted. The proposed DUST model has a function for exposure by direct spray and a function for exposure from foliage (indirect exposure).  

Collected amphibian dermal exposure data is from 9 studies. Van Meter et al. 2014/2015 data includes exposure with 5 active ingredients (Imidacloprid, Pendimethalin, Atrazine, Fipronil, and Triadimefon) and 9 amphibian species in the terrestrial metamorph stage (American toad, Barking treefrog, Cricket frog, Fowler's toad, Gray treefrog, Green treefrog, Leopard frog, Mole salamander, Narrowmouth toad). Van Meter et al. 2016 data includes exposure to the same 5 pesticides on American toads. Van Meter et al. 2018 data consisted of exposure to a single pesticide or pesticide mixtures (Atrazine, Metolachlor, 2,4-D, Malathion, Propiconazole) on juvenile green frogs. Glinski et al. 2018a data includes exposure to 5 pesticides (Atrazine, Triadimefon, Metolachlor, Chlorothalonil, Imidacloprid) on 2 amphibian species (Southern leopard frogs, Fowler's toads). Glinski et al. 2018b data consists of exposure to 3 pesticides (Atrazine, Triadimefon, Fopronil) on Fowler's toads. Glinski et al. 2018c data includes exposure to single, double, or triple pesticide mixtures (Bifenthrin, Metolachlor, Triadimefon) on Southern leopard frogs.  Glinski et al. 2020 data includes exposure to bifenthrin, chlorpyrifos, and trifloxystrobin on Leopard frogs. Henson-Ramsey 2008 data consists of exposure to malathion for Tiger salamanders.


```{r eval = TRUE, echo = FALSE, include = FALSE}
# Install required libraries
library(ggplot2)
require(gridExtra)
library(reshape2)

# tom epa linux
if(Sys.info()[4] == "d2626ut7920d.rtpnc.epa.gov"){
  amphibdir <- path.expand("~/git/amphib_dermal_collation/")
}
# Emma's EPA
if(Sys.info()[4]=="LZ2626UECHELSVI"){
  amphibdir <- path.expand("C:/Users/echelsvi/git/amphib_dermal_collation/")
}

# Variable names for data
amphibdir_data_in <- paste(amphibdir,'data_in/',sep='')
amphibdir_data_out <- paste(amphibdir,'data_out/',sep='')
amphibdir_graphics <- paste(amphibdir,'graphics/',sep='')
amphibdir_src <- paste(amphibdir,'src/',sep='')
```
<br><br>

***

**Collated Data Set**
============================================================================
**Data Set Dimensions, Column Names, and Summary:**
```{r eval = TRUE, echo = FALSE}
##################
#the data set
##################
# import data from consolidated database
combined_data_filename <- paste(amphibdir_data_out,"amphib_dermal_collated.csv", sep="")
combined_data <- read.csv(combined_data_filename)
dim(combined_data)
colnames(combined_data)
summary(combined_data)
```


~~ DELETE THIS ONCE GET APP RATES FOR DERMAL ROUTES DATA!!
```{r eval = TRUE, echo = TRUE}
combined_data <- combined_data[!(combined_data$source == "dag_dermal_routes"),]
```


~~ DELETE or PUT IN APPENDIX AT BOTTOM (unless decide it is useful elsewhere in the Rmd)  

Summary of app_rate       
```{r eval = TRUE, echo = FALSE, include = TRUE}
summary(combined_data$app_rate_g_cm2)
```

Summary of body_weight
```{r eval = TRUE, echo = FALSE, include = TRUE}
summary(combined_data$body_weight_g)
```

Summary of tissue_conc
```{r eval = TRUE, echo = FALSE, include = TRUE}
summary(combined_data$tissue_conc_ugg)
```

~~ DELETE or PUT IN APPENDIX AT BOTTOM (unless decide it is useful elsewhere in the Rmd)
```{r eval = TRUE, echo = FALSE, include = TRUE}
# App Rates
layout(matrix(c(1,2,3,4), nrow = 2, ncol = 2, byrow= TRUE)) #set the plotting area into a 2row*2col array
hist(combined_data$app_rate_g_cm2[combined_data$source == "rvm2015"], xlab = "Application Rate", main = "Van Meter 2014/15", col = "#feb24c")
hist(combined_data$app_rate_g_cm2[combined_data$source == "rvm2016"], xlab = "Application Rate", main = "Van Meter 2016", col = "#feb24c")
hist(combined_data$app_rate_g_cm2[combined_data$source == "rvm2017"], xlab = "Application Rate", main = "Van Meter 2017", col = "#feb24c")
hist(combined_data$app_rate_g_cm2[combined_data$source == "hr2008"], xlab = "Application Rate", main = "Henson-Ramsey 2008", col = "#feb24c")


layout(matrix(c(1,2,3,3), nrow = 2, ncol = 2, byrow= TRUE)) #set the plotting area into a 2row*2col array
hist(combined_data$app_rate_g_cm2[combined_data$source == "dag_dehydration"], xlab = "Application Rate", main = "Glinski (Dehydration)", col = "#feb24c")
hist(combined_data$app_rate_g_cm2[combined_data$source == "dag_metabolites"], xlab = "Application Rate", main = "Glinski (Metabolites)", col = "#feb24c")
hist(combined_data$app_rate_g_cm2[combined_data$source == "dag_biomarker"], xlab = "Application Rate", main = "Glinski (Biomarkers)", col = "#feb24c")
#hist(combined_data$app_rate_g_cm2[combined_data$source == "dag_dermal_routes"], xlab = "Application Rate", main = "Glinski (Dermal Routes)", col = "#feb24c")


# Body Weight
layout(matrix(c(1,2,3,4), nrow = 2, ncol = 2, byrow= TRUE)) #set the plotting area into a 2row*2col array
hist(combined_data$body_weight_g[combined_data$source == "rvm2015"], xlab = "Body Weight", main = "Van Meter 2014/15", col = "#969696")
hist(combined_data$body_weight_g[combined_data$source == "rvm2016"], xlab = "Body Weight", main = "Van Meter 2016", col = "#969696")
hist(combined_data$body_weight_g[combined_data$source == "rvm2017"], xlab = "Body Weight", main = "Van Meter 2017", col = "#969696")
hist(combined_data$body_weight_g[combined_data$source == "hr2008"], xlab = "Body Weight", main = "Henson-Ramsey 2008", col = "#969696")


layout(matrix(c(1,2,3,3), nrow = 2, ncol = 2, byrow= TRUE)) #set the plotting area into a 2row*2col array
hist(combined_data$body_weight_g[combined_data$source == "dag_dehydration"], xlab = "Body Weight", main = "Glinski (Dehydration)", col = "#969696")
hist(combined_data$body_weight_g[combined_data$source == "dag_metabolites"], xlab = "Body Weight", main = "Glinski (Metabolites)", col = "#969696")
hist(combined_data$body_weight_g[combined_data$source == "dag_biomarker"], xlab = "Body Weight", main = "Glinski (Biomarkers)", col = "#969696")
#hist(combined_data$body_weight_g[combined_data$source == "dag_dermal_routes"], xlab = "Body Weight", main = "Glinski (Dermal Routes)", col = "#969696")

# Tissue Concentration
layout(matrix(c(1,2,3,4), nrow = 2, ncol = 2, byrow= TRUE)) #set the plotting area into a 2row*2col array
hist(combined_data$tissue_conc_ugg[combined_data$source == "rvm2015"], xlab = "Tissue Concentration", main = "Van Meter 2014/15", col = "#fcc5c0")
hist(combined_data$tissue_conc_ugg[combined_data$source == "rvm2016"], xlab = "Tissue Concentration", main = "Van Meter 2016", col = "#fcc5c0")
hist(combined_data$tissue_conc_ugg[combined_data$source == "rvm2017"], xlab = "Tissue Concentration", main = "Van Meter 2017", col = "#fcc5c0")
hist(combined_data$tissue_conc_ugg[combined_data$source == "hr2008"], xlab = "Tissue Concentration", main = "Henson-Ramsey 2008", col = "#fcc5c0")


layout(matrix(c(1,2,3,3), nrow = 2, ncol = 2, byrow= TRUE)) #set the plotting area into a 2row*2col array
hist(combined_data$tissue_conc_ugg[combined_data$source == "dag_dehydration"], xlab = "Tissue Concentration", main = "Glinski (Dehydration)", col = "#fcc5c0")
hist(combined_data$tissue_conc_ugg[combined_data$source == "dag_metabolites"], xlab = "Tissue Concentration", main = "Glinski (Metabolites)", col = "#fcc5c0")
hist(combined_data$tissue_conc_ugg[combined_data$source == "dag_biomarker"], xlab = "Tissue Concentration", main = "Glinski (Biomarkers)", col = "#fcc5c0")
#hist(combined_data$tissue_conc_ugg[combined_data$source == "dag_dermal_routes"], xlab = "Tissue Concentration", main = "Glinski (Dermal Routes)", col = "#fcc5c0")
```


<br><br>

***

**TIM (Terrestrial Investigation Model)**
=====================================================================
By default, TIM estimates exposures for birds. The dermal exposure dose (body burden) is computed by combining the pesticide application rate (app_rate), surface area (sa_default_tim), fraction of surface area exposed (sa_tim_default_frac), body weight of the exposed organism (body_weight) and the dermal absorption factor (dermal_af).  


$$TIM = \frac{AR * SA * SAF * DAF}{BW}$$

```{r eval = TRUE, echo = FALSE, include = FALSE}
colnames(combined_data)
n <- length(combined_data$tissue_conc_ugg)

# bird model with perez body weight
#   we dont care about perez body weights because they are based on large pigeons
# sa_bird_perez <- 65.15 * (body_weight ^ 0.559)
```

***

### **TIM Default Calculation**
*Direct Exposure*
```{r eval = TRUE, echo = TRUE}
# Equation 1: TIM Default Values 
app_rate <- combined_data$app_rate_g_cm2
body_weight <- combined_data$body_weight_g
conv_rate <- 1000000.0
dermal_af <- 1.0
hours_one <- 1.0
sa_default_tim <- 10 * (body_weight ^ 0.667) 
sa_tim_default_frac <- 0.5
tissue_conc <- combined_data$tissue_conc_ugg
```

```{r eval = TRUE, echo = TRUE, include = TRUE}
# Calculate Equation 1a: TIM Default
combined_data$tim_direct_default <- (app_rate * conv_rate * sa_default_tim * sa_tim_default_frac * dermal_af)/body_weight
tim_direct_default <- combined_data$tim_direct_default
```

```{r eval = TRUE, echo = FALSE, include = TRUE}
hist(tim_direct_default, main = "TIM Default (Direct)", col = "#7a0177")
```


~~~~ MAYBE DELETE THIS OR PUT IN APPENDIX AT BOTTOM UNLESS USEFUL HERE/ELSEWHERE
```{r eval = TRUE, echo = FALSE, include = TRUE}
layout(matrix(c(1,2,3,4), nrow = 2, ncol = 2, byrow= TRUE)) #set the plotting area into a 2row*2col array
hist(combined_data$tim_direct_default[combined_data$source == "rvm2015"], main = "TIM Default (Direct)", xlab = "Van Meter 2014/15", col = "#7a0177")
hist(combined_data$tim_direct_default[combined_data$source == "rvm2016"], main = "TIM Default (Direct)", xlab = "Van Meter 2016", col = "#7a0177")
hist(combined_data$tim_direct_default[combined_data$source == "rvm2017"], main = "TIM Default (Direct)", xlab = "Van Meter 2017", col = "#7a0177")
hist(combined_data$tim_direct_default[combined_data$source == "hr2008"], main = "TIM Default (Direct)", xlab = "Henson-Ramsey 2008", col = "#7a0177")


layout(matrix(c(1,2,3,3), nrow = 2, ncol = 2, byrow= TRUE)) #set the plotting area into a 2row*2col array
hist(combined_data$tim_direct_default[combined_data$source == "dag_dehydration"], main = "TIM Default (Direct)", xlab = "Glinski (Dehydration)", col = "#7a0177")
hist(combined_data$tim_direct_default[combined_data$source == "dag_metabolites"], main = "TIM Default (Direct)", xlab = "Glinski (Metabolites)", col = "#7a0177")
hist(combined_data$tim_direct_default[combined_data$source == "dag_biomarker"], main = "TIM Default (Direct)", xlab = "Glinski (Biomarkers)", col = "#7a0177")
#hist(combined_data$tim_direct_default[combined_data$source == "dag_dermal_routes"], main = "TIM Default (Direct)", xlab = "Glinski (Dermal Routes)", col = "#7a0177")
```


### **TIM Default Ratios**
~~ blurb about how we can assess the ratios...1:1 would be perfect agreement
```{r eval = TRUE, echo = TRUE}
# Calculate Equation 1b: TIM Default Ratios
combined_data$tim_direct_default_ratios <- combined_data$tim_direct_default/tissue_conc
```

~~ talk about this range in values, and talk about the histogram (is the histogram helpful?  scatterplot better?)
```{r eval = TRUE, echo = TRUE, include = TRUE}
# Range
print(paste(min(combined_data$tim_direct_default_ratios), "-",   max(combined_data$tim_direct_default_ratios)))
```

```{r eval = TRUE, echo = FALSE, include = TRUE}
hist(combined_data$tim_direct_default_ratios, main = "Histogram of TIM Default Ratios", col = "#2b8cbe", xlab = "TIM Default Ratios")
```


~~~~ MAYBE DELETE THIS OR PUT IN APPENDIX AT BOTTOM UNLESS USEFUL HERE/ELSEWHERE (think scatterplots are probably better?)
```{r eval = TRUE, echo = FALSE}
# TIM Default Ratios
layout(matrix(c(1,2,3,4), nrow = 2, ncol = 2, byrow= TRUE)) #set the plotting area into a 2row*2col array
hist(combined_data$tim_direct_default_ratios[combined_data$source == "rvm2015"], main = "TIM Default Ratios", xlab = "Van Meter 2014/15", col = "#2b8cbe")
hist(combined_data$tim_direct_default_ratios[combined_data$source == "rvm2016"], main = "TIM Default Ratios", xlab = "Van Meter 2016", col = "#2b8cbe")
hist(combined_data$tim_direct_default_ratios[combined_data$source == "rvm2017"], main = "TIM Default Ratios", xlab = "Van Meter 2017", col = "#2b8cbe")
hist(combined_data$tim_direct_default_ratios[combined_data$source == "hr2008"], main = "TIM Default Ratios", xlab = "Henson-Ramsey 2008", col = "#2b8cbe")


layout(matrix(c(1,2,3,3), nrow = 2, ncol = 2, byrow= TRUE)) #set the plotting area into a 2row*2col array
hist(combined_data$tim_direct_default_ratios[combined_data$source == "dag_dehydration"], main = "TIM Default Ratios", xlab = "Glinski (Dehydration)", col = "#2b8cbe")
hist(combined_data$tim_direct_default_ratios[combined_data$source == "dag_metabolites"], main = "TIM Default Ratios", xlab = "Glinski (Metabolites)", col = "#2b8cbe")
hist(combined_data$tim_direct_default_ratios[combined_data$source == "dag_biomarker"], main = "TIM Default Ratios", xlab = "Glinski (Biomarkers)", col = "#2b8cbe")
#hist(combined_data$tim_direct_default_ratios[combined_data$source == "dag_dermal_routes"], main = "TIM Default Ratios", xlab = "Glinski (Dermal Routes)", col = "#2b8cbe")
```


~~ Skewed-Right because of Biomarkers data set, or because of another cause? (DELETE MAYBE)
```{r eval = TRUE, echo = TRUE, include = TRUE}
# Tissue Concentrations >= 0.10
large_tissue <- combined_data[combined_data$tissue_conc_ugg >= 0.10, ]
# Tissue Concentrations < 0.10
small_tissue <- combined_data[combined_data$tissue_conc_ugg < 0.10, ]

# Compare histograms
layout(matrix(c(1,2), nrow = 1, ncol = 2, byrow= TRUE)) #set the plotting area into a 1row*2col array
hist(combined_data$tim_direct_default_ratios[combined_data$tissue_conc_ugg >=0.10], main = "TIM Default Ratios of Obs. with \n Tissue Conc>=.10", xlab = "")
hist(combined_data$tim_direct_default_ratios[combined_data$tissue_conc_ugg <0.10], main = "TIM Default Ratios of Obs. with \n Tissue Conc <0.10", xlab="")

```

~~ Maybe look into biomarkers data, might be bc of the max application rates?

#### **Scatterplots**
~~ maybe exclude all the above histograms and just include these scatterplots?
```{r eval = TRUE, echo = FALSE, include = TRUE}
# Scatterplots comparing Measured vs. Modeled (Tissue Conc vs. TIM Default)
layout(matrix(c(1,2,3,4), nrow = 2, ncol = 2, byrow= TRUE)) #set the plotting area into a 2row*2col array
plot(x = combined_data$tissue_conc_ugg[combined_data$source == "rvm2015"], y = combined_data$tim_direct_default[combined_data$source == "rvm2015"], main = "Van Meter 2014/15", xlab= "Measured Tissue Concentrations", ylab = "TIM Default", pch = 20, col = "#fc8d59")
plot(x = combined_data$tissue_conc_ugg[combined_data$source == "rvm2016"], y = combined_data$tim_direct_default[combined_data$source == "rvm2016"], main = "Van Meter 2016", xlab= "Measured Tissue Concentrations", ylab = "TIM Default", pch = 20, col = "#fc8d59")
plot(x = combined_data$tissue_conc_ugg[combined_data$source == "rvm2017"], y = combined_data$tim_direct_default[combined_data$source == "rvm2017"], main = "Van Meter 2017", xlab= "Measured Tissue Concentrations", ylab = "TIM Default", pch = 20, col = "#fc8d59")
plot(x = combined_data$tissue_conc_ugg[combined_data$source == "hr2008"], y = combined_data$tim_direct_default[combined_data$source == "hr2008"], main = "Henson-Ramsey 2008", xlab= "Measured Tissue Concentrations", ylab = "TIM Default", pch = 20, col = "#fc8d59")


layout(matrix(c(1,2,3,3), nrow = 2, ncol = 2, byrow= TRUE)) #set the plotting area into a 2row*2col array
plot(x = combined_data$tissue_conc_ugg[combined_data$source == "dag_dehydration"], y = combined_data$tim_direct_default[combined_data$source == "dag_dehydration"], main = "Glinski (Dehydration)", xlab= "Measured Tissue Concentrations", ylab = "TIM Default", pch = 20, col = "#fc8d59")
plot(x = combined_data$tissue_conc_ugg[combined_data$source == "dag_metabolites"], y = combined_data$tim_direct_default[combined_data$source == "dag_metabolites"], main = "Glinski (Metabolites)", xlab= "Measured Tissue Concentrations", ylab = "TIM Default", pch = 20, col = "#fc8d59")
plot(x = combined_data$tissue_conc_ugg[combined_data$source == "dag_biomarker"], y = combined_data$tim_direct_default[combined_data$source == "dag_biomarker"], main = "Glinski (Biomarkers)", xlab= "Measured Tissue Concentrations", ylab = "TIM Default", pch = 20, col = "#fc8d59")
#plot(x = combined_data$tissue_conc_ugg[combined_data$source == "dag_dermal_routes"], y = combined_data$tim_direct_default[combined_data$source == "dag_dermal_routes"], main = "Glinski (Dermal Routes)", xlab= "Measured Tissue Concentrations", ylab = "TIM Default", pch = 20, col = "#fc8d59")

```

***

### **TIM Amphibian Calculation**
*Direct Exposure*  

To gain a better understanding of pesticide risks on amphibians, the dermal exposure dose was also computed with a modified allometric relationship between body weight and surface area that is better suited for amphibians (Hutchison et al. 1968, EPA 2009).
```{r eval = TRUE, echo = TRUE, include = TRUE}
# Equation 2: TIM Amphibians value (Hutchinson body weight)
sa_amphib_hutchinson <- 1.131 * (body_weight ^ 0.579)
```

```{r eval = TRUE, echo = FALSE}
# sa_amphib_frac <- 2.0
```

```{r eval = TRUE, echo = TRUE, include = TRUE}
# Calculate Equation 2a: TIM Amphibian
combined_data$tim_direct_amphib <- (app_rate * conv_rate * sa_amphib_hutchinson * sa_tim_default_frac * dermal_af)/body_weight
```

### **TIM Amphibian Ratios**

```{r eval = TRUE, echo = TRUE, include = TRUE}
# Calculate Equation 2b: TIM Amphibian Ratios
combined_data$tim_direct_amphib_ratios <- combined_data$tim_direct_amphib/tissue_conc
```

Shown below, the range of the TIM Amphibian Ratios spans from 0.1064 to 113,033.7779. The histogram is skewed to the right. However, the range of TIM Amphibian Ratios is narrower than that of the TIM Default Ratios. 
```{r eval = TRUE, echo = TRUE, include = TRUE}
# Range
print(paste(min(combined_data$tim_direct_amphib_ratios), "-",   max(combined_data$tim_direct_amphib_ratios)))
```


```{r eval = TRUE, echo = FALSE, include = TRUE}
hist(combined_data$tim_direct_amphib_ratios, main = "Histogram of TIM Amphibian Ratios", col = "#31a354", xlab = "TIM Amphibian Ratios")
```


Likewise to the spread of the TIM Default Ratios, the spread of the TIM Amphibian Ratios can best be explained by the tissue concentration values. For example, lower tissue concentrations (approximately less than 0.10) yield higher ratio values.


### **Compare TIM Amphibian Ratios of all Data Sets**

#### **Histograms**
```{r eval = TRUE, echo = FALSE}
# TIM Default Ratios
layout(matrix(c(1,2,3,4), nrow = 2, ncol = 2, byrow= TRUE)) #set the plotting area into a 2row*2col array
hist(combined_data$tim_direct_amphib_ratios[combined_data$source == "rvm2015"], main = "TIM Amphibian Ratios", xlab = "Van Meter 2014/15", col = "#31a354")
hist(combined_data$tim_direct_amphib_ratios[combined_data$source == "rvm2016"], main = "TIM Amphibian Ratios", xlab = "Van Meter 2016", col = "#31a354")
hist(combined_data$tim_direct_amphib_ratios[combined_data$source == "rvm2017"], main = "TIM Amphibian Ratios", xlab = "Van Meter 2017", col = "#31a354")
hist(combined_data$tim_direct_amphib_ratios[combined_data$source == "hr2008"], main = "TIM Amphibian Ratios", xlab = "Henson-Ramsey 2008", col = "#31a354")


layout(matrix(c(1,2,3,3), nrow = 2, ncol = 2, byrow= TRUE)) #set the plotting area into a 2row*2col array
hist(combined_data$tim_direct_amphib_ratios[combined_data$source == "dag_dehydration"], main = "TIM Amphibian Ratios", xlab = "Glinski (Dehydration)", col = "#31a354")
hist(combined_data$tim_direct_amphib_ratios[combined_data$source == "dag_metabolites"], main = "TIM Amphibian Ratios", xlab = "Glinski (Metabolites)", col = "#31a354")
hist(combined_data$tim_direct_amphib_ratios[combined_data$source == "dag_biomarker"], main = "TIM Amphibian Ratios", xlab = "Glinski (Biomarkers)", col = "#31a354")
#hist(combined_data$tim_direct_amphib_ratios[combined_data$source == "dag_dermal_routes"], main = "TIM Amphibian Ratios", xlab = "Glinski (Dermal Routes)", col = "#31a354")
```


#### **Scatterplots**
```{r eval = TRUE, echo = FALSE, include = TRUE}
# Scatterplots comparing Measured vs. Modeled (Tissue Conc vs. TIM Amphibian)
layout(matrix(c(1,2,3,4), nrow = 2, ncol = 2, byrow= TRUE)) #set the plotting area into a 2row*2col array
plot(x = combined_data$tissue_conc_ugg[combined_data$source == "rvm2015"], y = combined_data$tim_direct_amphib[combined_data$source == "rvm2015"], main = "Van Meter 2014/15", xlab= "Measured Tissue Concentrations", ylab = "TIM Amphibian", pch = 20, col = "#31a354")
plot(x = combined_data$tissue_conc_ugg[combined_data$source == "rvm2016"], y = combined_data$tim_direct_amphib[combined_data$source == "rvm2016"], main = "Van Meter 2016", xlab= "Measured Tissue Concentrations", ylab = "TIM Amphibian", pch = 20, col = "#31a354")
plot(x = combined_data$tissue_conc_ugg[combined_data$source == "rvm2017"], y = combined_data$tim_direct_amphib[combined_data$source == "rvm2017"], main = "Van Meter 2017", xlab= "Measured Tissue Concentrations", ylab = "TIM Amphibian", pch = 20, col = "#31a354")
plot(x = combined_data$tissue_conc_ugg[combined_data$source == "hr2008"], y = combined_data$tim_direct_amphib[combined_data$source == "hr2008"], main = "Henson-Ramsey 2008", xlab= "Measured Tissue Concentrations", ylab = "TIM Amphibian", pch = 20, col = "#31a354")


layout(matrix(c(1,2,3,3), nrow = 2, ncol = 2, byrow= TRUE)) #set the plotting area into a 2row*2col array
plot(x = combined_data$tissue_conc_ugg[combined_data$source == "dag_dehydration"], y = combined_data$tim_direct_amphib[combined_data$source == "dag_dehydration"], main = "Glinski (Dehydration)", xlab= "Measured Tissue Concentrations", ylab = "TIM Amphibian", pch = 20, col = "#31a354")
plot(x = combined_data$tissue_conc_ugg[combined_data$source == "dag_metabolites"], y = combined_data$tim_direct_amphib[combined_data$source == "dag_metabolites"], main = "Glinski (Metabolites)", xlab= "Measured Tissue Concentrations", ylab = "TIM Amphibian", pch = 20, col = "#31a354")
plot(x = combined_data$tissue_conc_ugg[combined_data$source == "dag_biomarker"], y = combined_data$tim_direct_amphib[combined_data$source == "dag_biomarker"], main = "Glinski (Biomarkers)", xlab= "Measured Tissue Concentrations", ylab = "TIM Amphibian", pch = 20, col = "#31a354")
#plot(x = combined_data$tissue_conc_ugg[combined_data$source == "dag_dermal_routes"], y = combined_data$tim_direct_amphib[combined_data$source == "dag_dermal_routes"], main = "Glinski (Dermal Routes)", xlab= "Measured Tissue Concentrations", ylab = "TIM Amphibian", pch = 20, col = "#31a354")

```
```{r eval = TRUE, echo = TRUE, include = TRUE}
# Tissue Concentrations >= 0.10
large_tissue <- combined_data[combined_data$tissue_conc_ugg >= 0.10, ]
summary(large_tissue$tim_direct_amphib_ratios)


# Tissue Concentrations < 0.10
small_tissue <- combined_data[combined_data$tissue_conc_ugg < 0.10, ]
summary(small_tissue$tim_direct_amphib_ratios)
```


***

**Assessing Model Performance**
=======================================================
The TIM Default and TIM Amphibian models, described above, were compared to one another and to the post-exposure amphibian body burdens found in the collated data set in order to assess model performance.


```{r eval = TRUE, echo = TRUE, include = TRUE}
# Create df of ratios
model_log_ratios <- as.vector(c(log10(combined_data$tim_direct_default_ratios), 
                        y=log10(combined_data$tim_direct_amphib_ratios)))
model_factor <- as.vector(c(rep('tim_default_direct',n),rep('tim_amphib_direct',n)))
df.ratios <- data.frame(x=model_factor, y=model_log_ratios)

# Create df of tissue_conc and TIM Amphibian
df.tim_amphib <- data.frame(x=tissue_conc, y=combined_data$tim_direct_amphib)
```

***

### **Comparing Models**
```{r eval = TRUE, echo = FALSE, include = TRUE}
ggplot(df.ratios, aes(x=x, y=y)) + geom_point() + labs(title="TIM Amphibian vs. TIM Default",  x="Model Factor", y = "TIM Value")

```

***

### **Plotting Modeled vs. Measured**

A scatter plot was utilized to first assess TIM Amphibian performance. A model along the 1:1 line would display perfect agreement with the measured values.  


```{r eval = TRUE, echo = FALSE, include = TRUE}
ggplot(df.tim_amphib, aes(x=x, y=y)) + geom_point() + labs(title="Modeled vs. Measured",  x="Tissue Concentrations", y = "TIM Amphibian Values")
```

***

### **Plotting log10 Ratio of Modeled Burdens to Observed Burdens**

Box plots were utilized to further assess model performance. Well-performing models would minimize error, and therefore would have a higher percentage of ratios near zero (log10 of unity).  


```{r eval = TRUE, echo = FALSE, include = TRUE}
fill <- "#4271AE"
line <- "#1F3552"
ggplot(df.ratios, aes(x=x,y=y)) +
    geom_boxplot(fill = fill, colour = line) + # 
    scale_x_discrete(name = "Model") +
    scale_y_continuous(name = "log10 of Model-Observation Ratio") +
    theme_bw()
model_boxplot_filename <- paste(amphibdir_graphics,"model_boxplot.png",sep='')
ggsave(model_boxplot_filename, device="png", width=4, height=4)
```

***

### **False Negative Rates**

Summary statistics for false negative rates were also computed for both models. False negative rates are defined as an outcome where the modeled value is less than the measured value. This rate is based on a Type II Error in statistical hypothesis testing. False negative rates assess how often a model under-estimates the measurement of interest.

**TIM Default**

As displayed below, there are zero false negative rates for TIM Default. But,  115/798 (14.4%) of the measurements are within an order of magnitude of the modeled value. 
```{r eval = TRUE, echo = TRUE, include = TRUE}
# False Negative Rate: TIM Default
sum(tissue_conc > tim_direct_default)

# False Negative Rate (x10): TIM Default
sum(tissue_conc*10 > tim_direct_default)
```


```{r eval = TRUE, echo = FALSE, include = TRUE}
#plot(tissue_conc, combined_data$tim_direct_avian_1hr)
plot(tissue_conc, combined_data$tim_direct_default, xlab = "Tissue Concentration", ylab = "TIM Default Value", main = "TIM Default Value vs. Tissue Concentration")
df.tim_default <- data.frame(x=tissue_conc, y=tim_direct_default)
```

**TIM Amphibian**

There are 107/798 (13.4%) false negative screen results for TIM Amphibian, and 576/798 (72.2%) measurements within an order of magnitude of the modeled value. TIM Amphibian uses a less conservative but more representative surface area calculation for amphibians. 
```{r eval = TRUE, echo = TRUE, include = TRUE}
# False Negative Rate: TIM Amphibian
sum(tissue_conc > combined_data$tim_direct_amphib)

# False Negative Rate (x10): TIM Amphibian
sum(tissue_conc*10 > combined_data$tim_direct_amphib)
```

```{r eval = TRUE, echo = FALSE}
plot(tissue_conc, combined_data$tim_direct_amphib, xlab = "Tissue Concentration", ylab = "TIM Amphibian Value", main = "TIM Amphibian Value vs. Tissue Concentration")
```

***

### **Comparing Surface Area Calculations** 

While TIM Amphibian had a higher percentage of false negative screening values than TIM Default, the difference in screening values can be attributed to the surface area calculation. Amphibian surface area, as compared to the TIM Default surface area based on birds, is extrapolated from measured body weights. The comparison between surface area calculations is depicted below.    

```{r eval = TRUE, echo = FALSE}
min_bw <- min(combined_data$body_weight_g, na.rm=T)
max_bw <- max(combined_data$body_weight_g, na.rm=T)
range_bw <- seq(min_bw, max_bw, by = 0.01)
default_sa <- 10*range_bw^0.667
hutchison_sa <- 1.131 * range_bw^0.579
#max(default_sa/hutchison_sa)
sa_data <- data.frame(range_bw, default_sa, hutchison_sa)
ggplot(sa_data, aes(range_bw)) + 
  labs(x = "body weight (g)", y = "Surface Area (cm^2)",color = "Method") +
  geom_line(aes(y = default_sa, colour = "Default")) + 
  geom_line(aes(y = hutchison_sa, colour = "Hutchison")) +
  theme_bw() +
  theme(legend.position =c(0.2,0.8))
  
sa_comparison_filename <- paste(amphibdir_graphics,"sa_comparison.png",sep='')
ggsave(sa_comparison_filename, device="png", width=4, height=4)

```



Session Information
```{r eval = TRUE, echo = FALSE}
sessionInfo()
```