---
title: "Assessing DUST Model Performance for Amphibian Dermal Exposure?"
---


```{r eval=TRUE, echo=FALSE, include = FALSE}
# Install required libraries
library(ggplot2)
require(gridExtra)
library(reshape2)

# Emma's laptop
if(Sys.info()[4]=="LZ2626UECHELSVI"){
  amphibdir <- path.expand("C:/Users/echelsvi/OneDrive - Environmental Protection Agency (EPA)/Profile/Documents/amphib_dermal_collation/amphib_dermal_collation-master/")
}

# Variable names for data
amphibdir_data_in <- paste(amphibdir,'data_in/',sep='')
amphibdir_data_out <- paste(amphibdir,'data_out/',sep='')
amphibdir_graphics <- paste(amphibdir,'graphics/',sep='')
amphibdir_src <- paste(amphibdir,'src/',sep='')
```

```{r eval = TRUE, echo = FALSE, include = FALSE}
##################
#the data set
##################
# import data from consolidated database
combined_data_filename <- paste(amphibdir_data_out,"amphib_dermal_collated.csv", sep="")
combined_data <- read.csv(combined_data_filename)
dim(combined_data)
colnames(combined_data)
summary(combined_data)
```

***

**DUST Model Parameters** 
=========================================================================

DUST (Document 8/19/15) attempts to model dermal dose body burdens for amphibians from either direct spray or indirect spray. The function equations for DUST (Direct Spray) and DUST (Indirect Spray) are shown below.

***

### **Dermal Dose (Direct Spray)**  



$Dermal Dose = (\frac{Arate* 11.2 * SAtotal * 0.5 * DAF}{BW}) * Fred$  

* *Arate* = application rate (lb a.i./A)
* *11.2* = conversion from lb a.i./A to ug a.i./g-bw
* *SAtotal* = total surface area of organism
* *0.5* = Assumed that either the top half or the bottom half of the animal will be in contact with the ground or direct spray
* *DAF* = dermal adsorption fraction used to account for pesticide specific data that define a fraction of the pesticide mass present on the animal that is actually absorbed
* *BW* = body weight
* *Fred* = Dermal route equivalency factor applied to dermal exposure to derive an estimate of oral dose. (Assumed to be 1 for amphibians)
* *SAtotal* = Surface area total = $a3 * BW^{b3}$
* *a3* = 1.131
* *b3* = 0.579

The surface area formula used is the same used by Van Meter et al. 2014/2015. 

##### **Building the Dermal Exposure Function**
``` {r eval = TRUE, echo = TRUE}
# Convert from g/cm2 to lbs/Acre for input into dermal spray dose function
convert_app_rate <- function(app_rate){
  
  new_app_rate <- (app_rate * 89217.9122)
  
  return(new_app_rate)
}

# Function to calculate Dermal Direct Spray Dose
dermal_spray_dose <- function(Arate, SAtotal, DAF, BW, Fred){
  
  dsd1 <- (((Arate * 11.2) * (SAtotal * 0.5) * DAF / BW )* Fred)
  
  return (dsd1)
}

# Test function returns a value
convert_app_rate(8)
dermal_spray_dose(7,8,9,10,11)
```

### **Dermal Dose (Indirect Spray: Foliage Contact)**
 
$Dermal Dose = (\frac{Cplant * Fdfr * Rfoliar contact (8*SAtotal * 0.079 * 0.1)}{BW}) * Fred$  

* *Cplant* = Concentration of the pesticide in crop foliage at time t. Residue value used for broadleaf foliage concentration in the assessment of dietary exposure
* *Cplant (for broadleaf plants)* = 135 * application rate (lbs a.i./hectare)
* *Fdfr* = Dislodgeable foliar residue adjustment factor
* *Rfoliar contact* = Rate of foliar contact
* *8* = number of hours per day the animal is active
* *SAtotal* = surface area total of the organism
* *0.079* = represents the fraction of the animal in contact with the foliage
* *0.1* = unit conversion to ug a.i./g-bw
* *BW* = body weight of organism
* *Fred* = Dermal route equivalency factor (Assumed to be 1 for amphibians)  

##### **Building the Dermal Exposure Function**
``` {r eval = TRUE, echo = TRUE}
# Function to calculate Cplant (broadleaf foliar residue from T-Rex)
Cplant_calc <- function(app_rate_g_cm2){
  
  Cplant_out <- (app_rate_g_cm2 * 220462 * 135)
    # 220462 = unit conversion from g/cm2 to lbs/hectare
    # 135 = used in T-Rex to calculate residue on broadleaf plant
  
  return(Cplant_out)
}

# Function to calculate Dermal Exposure (Indirect Foliar Contact)
dermal_contact_dose<-function(Cplant, Fdfr, Rfoliarcontact, SAtotal, BW, Fred){
  
  dcd1<-( Cplant * Fdfr * Rfoliarcontact * (8*SAtotal*0.079*0.1) / BW ) * Fred
  
  return (dcd1)
}

# Test function returns a value
out1 <- Cplant_calc(4)
dermal_contact_dose(out1,2,3,4,5,6)
```

**Assessing DUST Model Performance**
============================================================================

Using the input values from the Van Meter et al. database, the estimated dermal spray dose (direct spray) is computed. Later, it will be compared with the measured body burdens (tissue concentrations) found in the collated data set.

``` {r eval = TRUE, echo = TRUE}
# Surface Area calculation
combined_data$sa <- 1.131 * (combined_data$body_weight_g ^ 0.579)

# Estimated Dermal Dose (Direct Spray)
DAF <- 1 #constant
Fred <- 1 #constant
combined_data$app_rate_lbs_A <- convert_app_rate(combined_data$app_rate_g_cm)

combined_data$dermal_spray_dose_out <- dermal_spray_dose(combined_data$app_rate_lbs_A, combined_data$sa, DAF, combined_data$body_weight_g, Fred)

combined_data$ratio_out <- combined_data$dermal_spray_dose_out / combined_data$tissue_conc_ugg
```

The dermal dose (indirect spray) is also estimated with the Van Meter et al. database input values.  It will also be compared with the measured body burdens.   

``` {r eval = TRUE, echo = TRUE}
# Surface Area Calculation
combined_data$sa <- 1.131 * (combined_data$body_weight_g ^ 0.579)
DAF <- 1 #constant
Fred <- 1 #constant
Fdfr <- 0.62
Rfoliarcontact <- 6.01

combined_data$Cplant <- Cplant_calc(combined_data$app_rate_g_cm2)

combined_data$dermal_foliage_dose_out <- dermal_contact_dose(combined_data$Cplant, Fdfr, Rfoliarcontact, combined_data$sa, combined_data$body_weight_g, Fred)

combined_data$ratio_out_foliage <- combined_data$dermal_foliage_dose_out / combined_data$tissue_conc_ugg
```

***

### **Comparing Modeled vs. Measured**  

#### **DUST (Direct Spray)**
Fig. 1A compares the DUST (Direct Spray) equation modeled tissue concentrations with all the measured data. Fig. 1B is a histogram of the ratio of the DUST (Direct Spray) modeled output to all the measured tissue concentrations including direct and indirect exposure. Fig. 1A and 1B demonstrate the DUST (Direct Spray) model is non-conservative (that is, it under-estimates the body burden from direct spray) for a number of test cases, which we break out by chemicals later.

```{r fig.width = 8, fig.height = 6, echo = FALSE}
# Plots for Dermal Dose (Direct Spray) Equation with all body burden data
par(mfrow=(c(1,2)))
max1 <- max(combined_data$tissue_conc_ugg, na.rm=TRUE)
min <- 0
plot(combined_data$dermal_spray_dose_out ~ combined_data$tissue_conc_ugg, xlim = c(min, max1), ylim = c(min, max1), xlab = "Measured Body Burden", ylab = "DUST (Direct Spray)")
abline(0, 1, col = "red")
title(main = c("Figure 1"))
mtext("A", side=3, line=-1.5, at=par("usr")[1]+0.075*diff(par("usr")[1:2]), cex=1.2)
# Histogram of ratios
hist(combined_data$ratio_out, breaks = 50, xlab=c("Ratio of DUST to Measured"), main=c(""), col = "grey")
mtext("B",side=3,line=-1.5,at=par("usr")[1]+0.1*diff(par("usr")[1:2]), cex=1.2)
```

#### **DUST (Indirect Spray)**


Fig. 2A compares the DUST (Indirect Spray) equation modeled tissue concentrations with all of the measured data. Fig. 2B is a histogram of the ratio of the DUST (Indirect Spray) modeled output to all the measured tissue concentrations including direct and indirect exposure. Fig 2A and 2B demonstrate the DUST (Indirect Spray) model is conservative for all the test cases, unlike the DUST (Direct Spray) exposure model.  


~~~  This is extremely conservative, log scale? ~~~~


```{r fig.width = 8, fig.height = 6, echo = FALSE}
# Plots for Dermal Dose (Indirect Foliage) with all body burden data
par(mfrow=(c(1,2)))
max1 <- max(combined_data$dermal_foliage_dose_out, na.rm=TRUE)
min <- 0
plot(combined_data$dermal_foliage_dose_out ~ combined_data$tissue_conc_ugg, xlim = c(min,max1), ylim = c(min,max1), xlab = "Measured Body Burden", ylab = "DUST (Indirect Foliage)")
abline(0, 1, col="red")
title(main=c("Fig. 2"))
mtext("A",side=3,line=-1.5,at=par("usr")[1]+0.075*diff(par("usr")[1:2]), cex=1.2)
hist(combined_data$ratio_out_foliage, breaks = 50, xlab=c("Ratio of DUST to Measured"), main=c(""), col = "grey")
mtext("B",side=3,line=-1.5,at=par("usr")[1]+0.075*diff(par("usr")[1:2]), cex=1.2)
```



***


### **Breakdown of Model Performance by Chemical**

The DUST Model performance was assessed by examining body burdens for each chemical. To do this, a sub data set consisting only of the total active ingredients (imidacloprid, pendimethalin, total atrazine, total fipronil, total triadimefon) was created from the collated data set. 

```{r fig.width = 7, fig.height = 6, echo = FALSE, include = FALSE}
# creating consistent chemical names in dataset
unique.chemicals <- unique(combined_data$chemical)
# create consistent chemical names
combined_data$chemical[combined_data$chemical=="ATZ"] <-"Atrazine"
combined_data$chemical[combined_data$chemical=="Fip"] <-"Fipronil"
combined_data$chemical[combined_data$chemical=="TDN"] <-"Triadimefon"
combined_data$chemical[combined_data$chemical=="Pendi"] <-"Pendimethalin" #no metabolites measured
combined_data$chemical[combined_data$chemical=="Imid"] <-"Imidacloprid" #no metabolites measured
combined_data$chemical[combined_data$chemical=="ATZTOT"] <-"Total Atrazine"
combined_data$chemical[combined_data$chemical=="FipTOT"] <-"Total Fipronil"
combined_data$chemical[combined_data$chemical=="TNDTOT "] <-"Total Triadimefon"

chemicals2 <- c("Total Atrazine","Imidacloprid","Total Fipronil","Pendimethalin","Total Triadimefon")

# create dataframe with total active ingredients for use in comparison
combined_data2 <- subset(combined_data, chemical %in% c("Total Atrazine","Imidacloprid","Total Fipronil","Pendimethalin","Total Triadimefon"))
View(combined_data2)
combined_data2$chemical <- factor(combined_data2$chemical)
```

```{r eval = TRUE, echo = TRUE, include = TRUE}
dim(combined_data2)
summary(combined_data2$chemical)
```


Plots were created to compare modeled and measured data (direct and indirect) with total active ingredient data, colored by active ingredient.  


Figure 3A compares DUST (Direct Spray) modeled values with the measured tissue concentration from direct spray of five active ingredients. Figure 3B compares DUST (Direct Spray) modeled values with the measured tissue concentrations from indirect spray.   

~~ need to alter .. The model under-estimates body burdens for atrazine, triadimefon and fipronil exposures for measured tissue concentrations of amphibians exposed directly and indirectly. While the tissue concentrations from indirect uptake is less than tissue concentrations from indirect uptake, the model under-estimates body burdens for atrazine, fipronil, and triadimefon active ingredients. Pesticide concentration in an amphibian is clustered by active ingredient. There is no chemical property information mapped to bioconcentration factors in the current model that might help to deal with this clustering. Adding some chemical property information to the DUST model could be one way to deal with this clustering.

```{r eval = TRUE, fig.width = 8, fig.height = 6, echo = FALSE, include = TRUE}
par(mfrow=c(1,2))
# DUST (Direct Spray) vs. Measured (Direct Spray)
max1 <- max(combined_data2$tissue_conc_ugg)+1
min <- 0 
plot(combined_data2$dermal_spray_dose_out[which(combined_data2$application=="Direct")]~combined_data2$tissue_conc_ugg[which(combined_data2$application=="Direct")], xlab=c('Measured Tissue Concentration Direct'), ylab=c("Modeled Tissue Concentration Direct"), pty='s', xlim=c(min,max1), ylim=c(min,max1),pch=16, col=combined_data2$chemical[which(combined_data2$application=="Direct")])
abline(0, 1, col="red")
title(main=c("Fig. 3 DUST (Direct Spray)"))
#legend(x="topright", legend = levels(combined_data2$Chemical), col=1:nlevels(combined_data2$Chemical), pch=16)
mtext("A",side=3,line=-1.5,at=par("usr")[1]+0.075*diff(par("usr")[1:2]), cex=1.2)

# DUST (Direct Spray) vs. Measured (Indirect Spray)
max1<-max(combined_data2$tissue_conc_ugg)+1
min<-0 
plot(combined_data2$dermal_spray_dose_out[which(combined_data2$application=="indirect")]~combined_data2$tissue_conc_ugg[which(combined_data2$application=="indirect")], xlab=c('Measured Tissue Concentration Indirect'), ylab=c("Modeled Tissue Concentration Direct"),pty='s', xlim=c(min,max1), ylim=c(min,max1), pch=16,col=combined_data2$chemical[which(combined_data2$application=="indirect")] )
abline(0, 1, col="red")
title(main=c(" "))
legend(x="topright", legend = levels(combined_data2$chemical), col=1:nlevels(combined_data2$chemical), pch=16)
mtext("B",side=3,line=-1.5,at=par("usr")[1]+0.075*diff(par("usr")[1:2]), cex=1.2)

```

Figure 4A compares DUST (Indirect Spray) modeled values with the measured indirect uptake for five active ingredients. Figure 4B compares DUST (Indirect Spray) modeled values with the measured direct uptake for five active ingredients.   

The modeled values are conservative for all the active ingredients when comparing DUST modeled indirect foliage uptake values to either directly applied or indirectly applied body burdens.


```{r eval = TRUE, fig.width = 8, fig.height = 6, echo = FALSE, include = TRUE}
par(mfrow=c(1,2))
# indirect dermal DUST vs. measured indirect
max1 <- max(combined_data2$dermal_foliage_dose_out[which(combined_data2$application=="indirect")], na.rm = TRUE)+1
min <- 0 
plot(combined_data2$dermal_foliage_dose_out[which(combined_data2$application=="indirect")]~combined_data2$tissue_conc_ugg[which(combined_data2$application=="indirect")], xlab=c('Measured Tissue Concentration Indirect'), ylab=c("Modeled Tissue Concentration Indirect"),pty='s', xlim=c(min,max1), ylim=c(min,max1), pch=16,col=combined_data2$chemical[which(combined_data2$application=="indirect")])
abline(0, 1, col="red")
title(main=c("Fig. 4 DUST (Indirect Spray)"))
#legend(x="topright", legend = levels(combined_data2$Chemical), col=1:nlevels(combined_data2$Chemical), pch=16)
mtext("A",side=3,line=-1.5,at=par("usr")[1]+0.075*diff(par("usr")[1:2]), cex=1.2)

# DUST (Indirect Spray) vs. Measured (Direct Spray)
max1<-max(combined_data2$dermal_foliage_dose_out[which(combined_data2$application=="indirect")], na.rm=T)+1
min<-0 
plot(combined_data2$dermal_foliage_dose_out[which(combined_data2$application=="Direct")]~combined_data2$tissue_conc_ugg[which(combined_data2$application=="Direct")], xlab=c('Measured Tissue Concentration Direct'), ylab=c("Modeled Tissue Concentration Indirect"),pty='s', xlim=c(min,max1), ylim=c(min,max1), pch=16,col=combined_data2$chemical[which(combined_data2$application=="Direct")])
abline(0, 1, col="red")
title(main=c(" "))
legend(x="topright", legend = levels(combined_data2$chemical), col=1:nlevels(combined_data2$chemical), pch=16)
mtext("B",side=3,line=-1.5,at=par("usr")[1]+0.075*diff(par("usr")[1:2]), cex=1.2)

```


Fig. 5A and 5B are the same plots as Fig. 4 but with a log scale of the x and y axis in order to display the trends more clearly.

```{r eval = TRUE, fig.width=8, fig.height=6, echo = FALSE, include = TRUE}
par(mfrow=c(1,2))
# DUST (Indirect Spray) vs. Measured (Indirect Spray)
max1 <- max(combined_data2$dermal_foliage_dose_out[which(combined_data2$application=="indirect")], na.rm=T)+1
min <- 0 
plot(log(combined_data2$dermal_foliage_dose_out[which(combined_data2$application=="indirect")])~log(combined_data2$tissue_conc_ugg[which(combined_data2$application=="indirect")]), xlab=c('log(Measured Tissue Concentration Indirect)'), ylab=c("log(Modeled Tissue Concentration Indirect)"),pty='s', xlim=c(min,log(max1)), ylim=c(min,log(max1)), pch=16,col=combined_data2$chemical[which(combined_data2$application=="indirect")])
abline(0, 1, col="red")
title(main=c("Fig. 5 log(DUST Indirect Spray)"))
#legend(x="topright", legend = levels(combined_data2$Chemical), col=1:nlevels(combined_data2$Chemical), pch=16)
mtext("A",side=3,line=-1.5,at=par("usr")[1]+0.075*diff(par("usr")[1:2]), cex=1.2)

# indirect dermal DUST vs. measured direct
max1 <- max(combined_data2$dermal_foliage_dose_out[which(combined_data2$application=="indirect")], na.rm=T)+1
min <- 0 
plot(log(combined_data2$dermal_foliage_dose_out[which(combined_data2$application=="Direct")])~log(combined_data2$tissue_conc_ugg[which(combined_data2$application=="Direct")]), xlab=c('log(Measured Tissue Concentration Direct)'), ylab=c("log(Modeled Tissue Concentration Indirect)"),pty='s', xlim=c(min,log(max1)), ylim=c(min,log(max1)), pch=16,col=combined_data2$chemical[which(combined_data2$application=="Direct")])
abline(0, 1, col="red")
title(main=c(" "))
legend(x="topright", legend = levels(combined_data2$chemical), col=1:nlevels(combined_data2$chemical), pch=16)
mtext("B",side=3,line=-1.5,at=par("usr")[1]+0.075*diff(par("usr")[1:2]), cex=1.2)
```


Figure 6. The inputs are the combined data set of application rates and body weights from the amphibians exposed to derive body burdens. Indirect dermal uptake from foliage results are a magnitude of order higher than direct dermal spray uptake. 

Is it weird that indirect uptake has such higher uptake than direct input?

# Plots comparing modeled and measured data direct and indirect with total active ingredient data only colored by species.

```{r fig.width = 7, fig.height = 6}
# plot comparing direct vs. indirect DUST model output
plot(combined_data2$dermal_foliage_dose_out~combined_data2$dermal_spray_dose_out, col=combined_data2$chemical, xlab=c("DUST Direct"), ylab=c("DUST Indirect"), pch=16)
legend(x="topleft", legend = levels(combined_data2$chemical), col=1:nlevels(combined_data2$chemical), pch=16)
title(main=c("Fig. 6. Indirect DUST output vs. Direct DUST output"))
abline(0,1, col="red")
```



```{r fig.width = 8, fig.height = 6}
par(mfrow=c(1,2))
# direct dermal DUST vs. measured direct
max1 <- max(combined_data2$tissue_conc_ugg)+1
min <- 0 
plot(combined_data2$dermal_spray_dose_out[which(combined_data2$application=="Direct")]~combined_data2$tissue_conc_ugg[which(combined_data2$application=="Direct")], xlab=c('Measured Tissue Concentration Direct'), ylab=c("Modeled Tissue Concentration Direct"), pty='s', xlim=c(0,max1), ylim=c(0,max1),pch=16, col=combined_data2$species[which(combined_data2$application=="Direct")])
abline(0, 1, col="red")
title(main=c("Fig. 7 Direct DUST"))
#legend(x="topright", legend = levels(combined_data2$Species), col=1:nlevels(combined_data2$Species), pch=16)
mtext("A",side=3,line=-1.5,at=par("usr")[1]+0.075*diff(par("usr")[1:2]), cex=1.2)

# DUST (Direct Spray) vs. Measured Indirect
max1 <- max(combined_data2$tissue_conc_ugg)+1
min <- 0 
plot(combined_data2$dermal_spray_dose_out[which(combined_data2$application=="indirect")]~combined_data2$tissue_conc_ugg[which(combined_data2$application=="indirect")], xlab=c('Measured Tissue Concentration Indirect'), ylab=c("Modeled Tissue Concentration Direct"),pty='s', xlim=c(min,max1), ylim=c(min,max1), pch=16,col=combined_data2$species[which(combined_data2$application=="indirect")])
abline(0, 1, col="red", xlim=c(min, max1), ylim=c(min, max1))
title(main=c(" "))
legend(x="topright", legend = levels(combined_data2$species), col=1:nlevels(combined_data2$species), pch=16)
mtext("B",side=3,line=-1.5,at=par("usr")[1]+0.075*diff(par("usr")[1:2]), cex=1.2)
```

Fig. 7A and 7B Modeled direct overspray of tissue concentration from DUST underestimates measured tissue concentration of some species body burdens when compared to measured direct and indirect exposure tissue concentrations. Species identity could play a role in pesticide uptake but clustering by active ingredient showed stronger clustering patterns. Accounting for species identity in uptake may not account for much variability in uptake as chemical property such as Kow, solubility, or molecular weight. 

```{r eval = FALSE, fig.width = 8, fig.height = 6}
par(mfrow=c(1,2))
# DUST (Indirect Spray) vs. Measured indirect
max1 <- max(combined_data2$TissueConc)+1
min <- 0 
plot(combined_data2$dermal_foliage_dose_out[which(combined_data2$application=="indirect")]~combined_data2$tissue_conc_ugg[which(combined_data2$application=="indirect")], xlab=c('Measured Tissue Concentration Indirect'), ylab=c("Modeled Tissue Concentration Indirect"),pty='s', xlim=c(min,max1), ylim=c(min,max1), pch=16,col=combined_data2$species[which(combined_data2$application=="indirect")])
abline(0, 1, col="red")
title(main=c("Fig. 8. Indirect DUST"))
#legend(x="topright", legend = levels(combined_data2$Species), col=1:nlevels(combined_data2$Species), pch=16)
mtext("A",side=3,line=-1.5,at=par("usr")[1]+0.075*diff(par("usr")[1:2]), cex=1.2)

# indirect dermal DUST vs. measured direct
max1<-max(combined_data2$tissue_conc_ugg[which(combined_data2$application=="Direct")])
min<-0 
plot(combined_data2$dermal_foliage_dose_out[which(combined_data2$application=="Direct")]~combined_data2$tissue_conc_ugg[which(combined_data2$application=="Direct")], xlab=c('Measured Tissue Concentration Direct'), ylab=c("Modeled Tissue Concentration Indirect"), pty='s', xlim=c(min,max1), ylim=c(min,max1), pch=16,col=combined_data2$species[which(combined_data2$application=="Direct")])
abline(0, 1, col="red")
title(main=c(" "))
legend(x="topright", legend = levels(combined_data2$species), col=1:nlevels(combined_data2$species), pch=16)
mtext("B",side=3,line=-1.5,at=par("usr")[1]+0.075*diff(par("usr")[1:2]), cex=1.2)
```

Fig. 8. Modeled indirect overspray of tissue concentration from DUST is conservative of measured tissue concentration of some species body burdens when compared to measured direct and indirect exposure body burdens. Fig. 8B which is Based on data from green treefrog and barking treefrog species the indirect DUST output is conservative of direct overspray measured body burdens.


Regression models with measured databased to determine if any of the pesticide chemical properties might be considered in the model as part of the bioconcentration factor. Pesticide properties considered are log Kow, molecular mass, and solubility i water. 


```{r eval = TRUE, echo = FALSE}
# add pesticide properties to combined_data2 table

pesticide_properties_file <- paste(amphibdir_data_in,"pesticide_properties.csv", sep="")
file.exists(pesticide_properties_file)
pesticide_properties <- read.table(pesticide_properties_file, header = TRUE, sep = ",")

combined_data3 <- merge(combined_data2, pesticide_properties, by.x="chemical", by.y="pesticide")

View(combined_data3)
```

```{r eval = FALSE, echo = FALSE}
library(MASS)
fit <- lm(y~x1+x2+x3,data = mydata)
step <- stepAIC(fit, direction="both")
step$anova # display results 

lm_all <- lm(tissue_conc~logkow+solubility+molecular_mass, data=combined_data3)
step <- stepAIC(lm_all, direction="both")
step$anova
```
